Viscous Transport in Eroding Porous Media

### 2019-10-01 docs/transport/erosion_transport.tex
small changes

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transport, Tracers, and Tortuosity}
\label{sec:transport}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Erosion in porous geometries leads to particular phenomena such as
channelization~\cite{berhanu2012shape}, and we are interested in
characterizing transport in such geometries.  In our previous
work~\cite{qua-moo2018}, we examined the effect of erosion on the area
fraction, flow rate, and the total drag.  However, to characterize
transport, other quantities must be examined.  Here, we compute the
anomalous dispersion rate, the tortuosity, and the distribution of the
pore sizes.  The first two metrics can be defined in terms of
streamlines governed by an advection autonomous
equation~\eqref{eqn:tracers}.

@@

We consider an eroding body that is initially a circle of radius
$r=0.4$.  We discretize the eroding body with $N_\iin=256$
discretization points, and its center is chosen so that the distance
between the body and the solid wall is $h/2$, $h/4$, and $h/10$, where
$h = 2\pi r/N_\iin$.  If we used the trapezoid rule and required an
error that is comparable to the Barycentric quadrature rule, the body
with an initial distance of $h/10$ from the solid wall would require
$6,400$ discretization points, and the outer wall would require $50,000$
discretization points.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Medium Porosity}
\label{sec:Eroding20}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding bodies that are discretized with $N_\iin=256$
points.  The background flow is the Hagen-Poiseuille flow
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right],
\end{align}
where the flow rate $U$ is chosen so that the pressure drop from $x=-2$
to $x=2$ is held fixed at 2.  Snapshots of the bodies at four equispaced
times are shown in Figure~\ref{fig:Eroding20vort}.  The color is the
vorticity, and the erosion rate is proportional to the vorticity.
Initially, several of the eroding bodies are closer to the outer wall
than the $5h$ threshold required to guarantee that the trapezoid rule
achieves machine precision.  In particular, the distance between bodies
1, 6, 13, and 15 and the outer wall is $1.3h$, $2.9h$, $2.8h$, and
$1.3h$, respectively, where $h$ is the arclength spacing of the outer
wall $\Gamma$.  In addition, the distance between several pairs of
eroding bodies, including 1 \& 6, 3 \& 9, 6 \& 8, and 14\& 18, is too
small to be accurately resolved with the trapezoid rule.  By using the
Barycentric quadrature rule, the interaction between these
nearly-touching bodies can be accurately resolved at the prescribed
resolution, and erosion can be simulated until all the bodies have
vanished.

@@

In Figure~\ref{fig:ErodingLow20Transport}(b), we plot the temporal
evolution of the particle spreading $\sigma_\lambda$.  As in the last
example, we analyze the spreading at several different porosities and we
use the reinsertion algorithm described in Section~\ref{sec:dispersion}
to form sufficiently long trajectories to observe an asymptotic
dispersion rate.  For all the porosities, the dispersion is much closer
to ballistic when compared to the results in
Figure~\ref{fig:Eroding20Transport}.  However, there are still clear
transitions from ballistic dynamics to asymptotic super-dispersive
spreading.  In contrast to the lower porosity initial condition
(Section~\ref{sec:Eroding20}), at early times the erosion results in a
decrease in the dispersion rate. In particular, once erosion has eroded
the first 5\% of the bodies, the particle spreading transitions from
$\sigma_\lambda \sim t^{0.92}$ to $\sigma_\lambda \sim t^{0.69}$.  To
explain this behavior, recall that anomalous dispersion is caused by
tracers spending time in both the fast and slow regimes.  Since the
initial configuration has a reasonably uniform velocity (see
Figure~\ref{fig:ErodingLow20vort}), albeit a small one, the dispersion
is nearly ballistic. However, as the geometry erodes, the flow becomes
more intermittent, and this results in an increased anomalous dispersion
rate~\cite{dea-leb-den-tar-bol-dav2013}.  Then, as the bodies continue
to erode, the geometry channelizes, and most tracers are transported
with a large velocity through the channels, again resulting in a nearly
ballistic motion~\cite{sie-ili-pri-riv-gua2019}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transport, Tracers, and Tortuosity}
\label{sec:transport}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Erosion in porous geometries leads to particular phenomena such as
channelization~\cite{berhanu2012shape}, and we are interested in
characterizing transport in such geometries.  In our previous
work~\cite{qua-moo2018}, we examined the effect of erosion on the area
fraction, flow rate, and the total drag.  However, to characterize
macroscopic signatures of the transport, other quantities must be
examined.  Here, we compute the anomalous dispersion rate, the
tortuosity, and the distribution of the pore sizes.  The first two
metrics can be defined in terms of streamlines governed by an advection
autonomous equation~\eqref{eqn:tracers}.

@@

We consider an eroding body that is initially a circle of radius
$r=0.4$.  We discretize the eroding body with $N_\iin=256$
discretization points, and its center is chosen so that the distance
between the body and the solid wall is $h$, $h/2$, and $h/10$, where $h
= 2\pi r/N_\iin$.  If we used the trapezoid rule and required an error
that is comparable to the Barycentric quadrature rule, the body with an
initial distance of $h/10$ from the solid wall would require $6,400$
discretization points, and the outer wall would require $50,000$
discretization points.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Medium Porosity}
\label{sec:Eroding20}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding bodies that are discretized with $N_\iin=256$
points.  The background flow is the Hagen-Poiseuille flow
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right],
\end{align}
where the flow rate $U$ is chosen so that the pressure drop from $x=-2$
to $x=2$ is held fixed at 8. Therefore, $U=1$ once all the bodies have
vanished.  Snapshots of the bodies at four equispaced times are shown in
Figure~\ref{fig:Eroding20vort}.  The color is the vorticity, and the
erosion rate is proportional to the vorticity.  Initially, several of
the eroding bodies are closer to the outer wall than the $5h$ threshold
required to guarantee that the trapezoid rule achieves machine
precision.  In particular, the distance between bodies 1, 6, 13, and 15
and the outer wall is $1.3h$, $2.9h$, $2.8h$, and $1.3h$, respectively,
where $h$ is the arclength spacing of the outer wall $\Gamma$.  In
addition, the distance between several pairs of eroding bodies,
including 1 \& 6, 3 \& 9, 6 \& 8, and 14\& 18, is too small to be
accurately resolved with the trapezoid rule.  By using the Barycentric
quadrature rule, the interaction between these nearly-touching bodies
can be accurately resolved at the prescribed resolution, and erosion can
be simulated until all the bodies have vanished.

@@

In Figure~\ref{fig:ErodingLow20Transport}(b), we plot the temporal
evolution of the particle spreading $\sigma_\lambda$.  As in the last
example, we analyze the spreading at several different porosities and we
use the reinsertion algorithm described in Section~\ref{sec:dispersion}
to form sufficiently long trajectories to observe an asymptotic
dispersion rate.  For all the porosities, the dispersion is much closer
to ballistic when compared to the results in
Figure~\ref{fig:Eroding20Transport}.  However, there are still clear
transitions from ballistic dynamics to asymptotic super-dispersive
spreading.  In contrast to the higher porosity initial condition
(Section~\ref{sec:Eroding20}), at early times the erosion results in a
decrease in the dispersion rate. In particular, once erosion has eroded
the first 5\% of the bodies, the particle spreading transitions from
$\sigma_\lambda \sim t^{0.92}$ to $\sigma_\lambda \sim t^{0.69}$.  To
explain this behavior, recall that anomalous dispersion is caused by
tracers spending time in both the fast and slow regimes.  Since the
initial configuration has a reasonably uniform velocity (see
Figure~\ref{fig:ErodingLow20vort}), albeit a small one, the dispersion
is nearly ballistic. However, as the geometry erodes, the flow becomes
more intermittent, and this results in an increased anomalous dispersion
rate~\cite{dea-leb-den-tar-bol-dav2013}.  Then, as the bodies continue
to erode, the geometry channelizes, and most tracers are transported
with a large velocity through the channels, again resulting in a nearly
ballistic motion~\cite{sie-ili-pri-riv-gua2019}.
